---
title: "Filter.Join"
author: "Abby Gillingham"
date: "2023-04-03"
output: html_document
---

```{r packages, include=F}
library(tidyverse) #install.packages("tidyvers")
library(readxl) #install.packages("readxl")
library(here) #install.packages("here")
library(kableExtra) #install.packages("kableExtra")
library(gapminder) #install.packages("gapminder")
library(hablar) #install.packages("hablar")
```

```{r read in files, include=F}
fish <- read_csv(here("data", "fish.csv"))
kelp_abur <- read_excel(here("data", "kelp.xlsx"), sheet="abur")
```

#### Filter rows by a common string:
```{r}
fish_garibaldi <- fish %>% filter(common_name == "garibaldi")
fish_garibaldi
```

#### Activity 8.3.1.1
```{r}
fish_mohk <- fish %>% filter(site == "mohk")
fish_mohk
```

#### Filter rows by numeric conditions:
```{r}
fish_over50 <- fish %>% filter(total_count >= 50)
fish_over50
```

#### Filter rows by multiple common strings:
```{r}
fish_3sp <- fish %>% filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))
fish_3sp
```

#### Activity 8.3.3.1
```{r}
fish_gar_2016 <- fish %>% filter(common_name=="garibaldi" | year=="2016")
fish_gar_2016
fish %>% filter(common_name=="garibaldi", year=="2016")
```

#### Activity 8.3.5
```{r}
low_gb_wr <- fish %>% filter(common_name=="garibaldi" | common_name=="rock wrasse") %>% filter(total_count <= 10)
low_gb_wr
```
Original solution:
```{r}
fish %>% filter(common_name %in% c("garibaldi", "rock wrasse"), total_count <= 10)
```

## Filtering with **AND** and **OR** at the same time:
#### Use "|" for OR, and "," for AND, then separate the OR and AND with separate "filter()" functions:
```{r}
fish %>% filter(common_name=="garibaldi" | common_name=="rock wrasse") %>% filter(total_count <= 10)
```


#### Filter by **detecting** a partial pattern:
```{r}
fish_black <- fish %>% filter(str_detect(common_name, pattern = "black"))
fish_black
```

#### Activity 8.3.7
```{r}
fish_it <- fish %>% filter(str_detect(common_name, pattern = "it"))
fish_it
```

### Joining:
#### Use "full_join()" when in doubt; it will combine everything from two datasets.
```{r}
abur_kelp_fish <- kelp_abur %>% full_join(fish, by=c("year", "site"))
abur_kelp_fish
```

#### Left joining with "left_join()":
```{r}
kelp_fish_left <- kelp_abur %>% left_join(fish, by=c("year", "site"))
kelp_fish_left
```
```{r right join, include=F}
kelp_fish_right <- kelp_abur %>% right_join(fish, by=c("year", "site"))
kelp_fish_right
```
Joining left or right maintains all columns from the table indicated (left table for left_join() and right table for right_join()) and adds data from the other table that matches with the first table.

#### Inner joining with "inner_join()":
```{r}
kelp_fish_injoin <- kelp_abur %>% inner_join(fish, by=c("year", "site"))
kelp_fish_injoin
```

#### Challenge 8.4.4
```{r}
fish_17_abur <- fish %>% filter(year=="2017", site=="abur")
kelp_abur1 <- kelp_abur %>% mutate(year=str_replace(year, pattern="17", replacement="2017")) %>% mutate(year=str_replace(year, pattern="18", replacement="2018")) %>% mutate(year=str_replace(year, pattern="19", replacement="2019")) %>% convert(num(year))

fish_17_abur_joined <- fish_17_abur %>% left_join(kelp_abur1, by=c("year", "site"))

my_fish_joined <- fish_17_abur_joined %>% mutate(fish_per_frond = total_count/total_fronds)
my_fish_joined
```

Also do the same thing more concisely (still have to re-format kelp_abur):
```{r}
my_fish_joined1 <- fish %>% filter(year==2017, site=="abur") %>% left_join(kelp_abur1, by=c("year", "site")) %>% mutate(fish_per_frond = total_count/total_fronds)
my_fish_joined1
```

```{r include=F}
kable(my_fish_joined)
my_fish_joined %>% kable() %>% kable_styling(bootstrap_options="striped", full_width=F)
```

```{r}
colnames(my_fish_joined) = c("Year", "Site", "Common Name", "Total", "Month", "Total Fronds", "Fish per Frond")
my_fish_joined
```

