---
title: "Graphing with R using ggplot2.03272024"
author: "Abby Gillingham"
date: "2023-03-27"
output: html_document
---

install.packages("tidyverse")
install.packages("readxl")
install.packages("here")
install.packages("readr")

library(tidyverse)
library(readxl)
library(here)
library(readr)

```{r CSV}
ca_np <- read_csv("ca_np.csv")
```

```{r EXCEL}
ci_np <- read_excel("ci_np.xlsx")
```

gg_base <- ggplot(data = ci_np, aes(x=year, y=visitors))
gg_base + geom_line()
gg_base + geom_point()
gg_base + geom_col()
gg_base + geom_area()
gg_base + geom_line(color = "purple", linetype="dashed")
gg_base + geom_point(color = "green", pch = 17, size = 4, alpha = 0.25)
gg_base + geom_area(color = "red", fill = "pink", alpha = 0.3)
gg_base + geom_point(aes(size = visitors, color = year), alpha = 0.5)
gg_base + geom_col(aes(fill = visitors))
gg_base + geom_point(aes(size = visitors, color = visitors), alpha = 0.4) + theme_bw()
gg_base + geom_line(linetype = "dotted", size = 1) + theme_bw() + labs(x = "Year", y = "Annual Park Visitors", title = "Channel Island NP Visitation", subtitle = "(1963-2016")
gg_base + geom_line(color = "red") + geom_point(color = "pink", aes(size = year))
gg_base + geom_col(fill = "pink", color = "red") + geom_line(color = "purple")

gg1_base <- ggplot(data = ca_np, aes(x=year, y=visitors, group = park_name)) 
gg1_base + geom_line(aes(color = park_name)) + labs(x="Year", y="Number of Visitors", title="National Park Visitors Per Year By Park", subtitle="(1900 - 2020)", col="National Park")

gg1_base + geom_line(show.legend = F, aes(color = park_name)) + labs(x="Year", y="Annual Visitors") + facet_wrap(~ park_name)


