---
title: "Pivot Lobsters"
author: "Abby Gillingham"
date: "2023-03-30"
output: html_document
---

```{r load-packages, include = F}
library(skimr) #install.packages("skimr")
library(kableExtra) #install.packages("kableExtra")
library(dplyr) #install.packages("dplyr")
library(magrittr) #install.packages("magrittr")
library(tidyverse)
library(readxl)
library(here)
options(dplyr.summarise.inform=F)
```


```{r read in data skip first four lines, include=F}
lobsters <- read_xlsx(here("data/lobsters.xlsx"), skip=4)
```

## Explore Data: Basic skim Function
```{r explore data}
skimr::skim(lobsters)
```

## Building Pivot Tables
Group by year, site, color, National Park, etc.
Summarize within that group using functions like count (n()), or mean (mean()) or standard deviation (sd()).
```{r}
lobsters %>% group_by(year) %>% summarize(count_by_year = n())
```

Group by multiple variables:
```{r}
lobsters %>% group_by(site, year) %>% summarize(count_by_siteyear=n())
```

And summarize multiple variables with multiple functions:
```{r}
siteyear_summary <- lobsters %>% group_by(site, year) %>% summarize(count_by_siteyear=n(), mean_size_mm=mean(size_mm), sd_size_mm=sd(size_mm))
siteyear_summary
```

## Customizing Pivot Tables:
* If some data is listed as "NA", you can remove this missing data using "na.rm=T", with na referring to the missing data, rm meaning remove, and T meaning T (i.e. "See that missing data? Remove it.)
```{r}
lobsters %>% group_by(site, year) %>% summarize(count_by_siteyear=n(), mean_size_mm=mean(size_mm, na.rm=T), sd_size_mm=sd(size_mm, na.rm=T))
```

* To make your pivot table headers prettier, assign names in quotation marks:
```{r}
siteyear_summary_labs <- lobsters %>% group_by("Site"=site, "Year"=year) %>% summarize("Count"=n(), "Mean"=mean(size_mm, na.rm=T), "St.Dev"=sd(size_mm, na.rm=T))
siteyear_summary_labs
```

* To format your table further, use kable():
```{r}
siteyear_summary_labs %>% kable()
```

### 6.4.6 Activity
```{r}
mean_by_siteyear_plot <- lobsters %>% group_by("Site"=site, "Year"=year) %>% summarize("Mean"=mean(size_mm, na.rm=T))
```

```{r plot mean size by site and year}
mean_by_siteyear <- ggplot(data=mean_by_siteyear_plot, aes(x=Year, y=Mean))
mean_by_siteyear+geom_col(aes(fill=Site))+labs(y="Mean Size (mm)", title="Mean Lobster Size by Site", subtitle="(2012 - 2016)", col="Site")+theme(axis.text.x=element_text(angle=90))+facet_wrap(~Site)
```

```{r}
mean_by_siteyear+geom_line(aes(color=Site))+labs(y="Mean Size (mm)", title="Mean Lobster Size by Site", subtitle="(2012 - 2016)", col="Site")
```

##### Real Solution:
```{r}
lobsters %>% group_by(site, year) %>% summarize(count_by_siteyear=n(), mean_size_mm=mean(size_mm, na.rm=T), sd_size_mm=sd(size_mm, na.rm=T), median_size_mm=median(size_mm, na.rm=T))
```

```{r}
lobsters %>% mutate(size_m=size_mm/1000)
lobsters_detailed <- lobsters %>% mutate(size_m=size_mm/1000, millenia=2000, observer="AH")
```

```{r}
lobsters_detailed %>% select(date, site, size_m)
```


