---
title: "Tidying"
author: "Abby Gillingham"
date: "2023-04-03"
output: html_document
---

```{r packages, include=F}
library(tidyverse) #install.packages("tidyverse")
library(here) #install.packages("here")
library(janitor) #install.packages("janitor")
library(readxl) #install.packages("readxl")
```

```{r read in inverts file, include=F}
inverts <- read_excel(here("data", "inverts.xlsx"))
view(inverts)
names(inverts)
summary(inverts)
```

#### Convert data that is too wide into a longer format:
```{r}
inverts_long <- pivot_longer(data=inverts, cols="2016":"2018", names_to="year", values_to="sp_count")
inverts_long
class(inverts_long)
```
#### Use "mutate()" to convert our new year column to numberic data:
```{r}
inverts_long <- inverts_long %>% mutate(year=as.numeric(year))
class(inverts_long$year)
```

#### Convert data that is too long to a wider format:
```{r}
inverts_wide <- inverts_long %>% pivot_wider(names_from=common_name, values_from=sp_count)
inverts_wide
```

#### Use "janitor::clean_names()" to automatically convert column names to lower_snake_case:
```{r}
inverts_wide <- inverts_wide %>% clean_names()
names(inverts_wide)
```

#### Combine, or "Unite" columns and name the new column:
```{r}
inverts_unite <- inverts_long %>% unite(col="site_year", c(site, year), sep="_")
inverts_unite
```

#### Activity 7.6.1.1
```{r}
inverts_moyr <- inverts_long %>% unite(col="mo_yr", c(month, year), sep="/")
inverts_moyr
```

#### Unite more than two columns:
```{r}
inverts_triple_unite <- inverts_long %>% unite(col="year_site_name", c(year, site, common_name), sep="-")
inverts_triple_unite
```

#### Separate a **character** column into multiple columns:
```{r}
inverts_sep <- inverts_unite %>% separate(site_year, into=c("my_site", "my_year"))
inverts_sep
```

#### Replace strings throughout a dataset with "str_replace()":
```{r}
ca_abbr <- inverts %>% mutate(common_name=str_replace(common_name, pattern="california", replacement="CA"))
ca_abbr
```

